---
title: "Using RUB colors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using RUB colors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(RUBer)
```

This vignette describes how to display and retrieve the RUB colors and palettes. For details on using these palettes in ggplot2, see the separate vignette XXX.

# Colors
`RUBer` provides easy access to the colors as specified in the [RUB corporate
design](https://www.ruhr-uni-bochum.de/cd/). Once the package is loaded, access 
all available colors by typing `RUB_colors`. Note that I assigned that the color
names were assigned as part of this package and are not offically sanctioned by
the corporate design guide.

```{r rub-colors}
# Retrieve all available colors
RUB_colors

# ... access individual color by index
RUB_colors[1]

# ... or by name
RUB_colors["dark red"]
```

To retrieve several colors at once, use the getter function 
`get_RUB_colors()`.

```{r get-rub-colors}
# Retrieve several colors by name
get_RUB_colors("green", "blue", "orange")

# Retrieve several colors by index
get_RUB_colors(4:6, 9)
```

# Palettes
The colors are combined in several palettes. To retrieve all available palettes,
use `RUB_palettes`. To retrieve a particular palette by name, use the
function `get_RUB_palettes` plus the number of required colors in round
brackets like so `get_RUB_palettes(palette = "discrete_5")(5)`. If the
palette contains less elements than requested, additional colors will
automatically be extrapolated.

Note that the RUB colors as defined by the corporate design were not designed
with data visualization in mind. The palettes presented here are ad-hoc and
rather informal, until the corporate design is offically extended to include
visualization design.

```{r get-rub-palette}

# All availabble palettes
RUB_palettes

# Retrieve five colors from the palette "discrete_5"
get_RUB_palettes(palette = "discrete_5")(5)

# Retrieve ten colors, five of which are extrapolated, from the palette 
# "discrete_5".
get_RUB_palettes(palette = "discrete_5")(10)
```

We can visualize all palettes using Emil Hvitfeldt's excellent `prismatic` 
package.


## Plots for the continuous palettes
First off, we have the two continuous palettes, "continuous" and 
"continuous_diverging".

```{r plot-continuous}
plot(prismatic::color(RUBer::get_RUB_palettes(palette = "continuous")(100)))

plot(prismatic::color(RUBer::get_RUB_palettes(palette = "continuous_diverging")(100)))
```

## Plots for the numbered discrete palettes
Second, we have all the numbered discrete palettes, which, for example, were
used to display the distribution of answers for the survey items. numbered
discrete palettes are pre-defined up to a total of eight separate colors, after,
it is necessary to rely on extrapolation.

```{r plot-discrete-numbered}
plot(prismatic::color(get_RUB_palettes(palette = "discrete_1")(1)))
plot(prismatic::color(get_RUB_palettes(palette = "discrete_2")(2)))
plot(prismatic::color(get_RUB_palettes(palette = "discrete_3")(3)))
plot(prismatic::color(get_RUB_palettes(palette = "discrete_4")(4)))
plot(prismatic::color(get_RUB_palettes(palette = "discrete_5")(5)))
plot(prismatic::color(get_RUB_palettes(palette = "discrete_6")(6)))
plot(prismatic::color(get_RUB_palettes(palette = "discrete_7")(7)))
plot(prismatic::color(get_RUB_palettes(palette = "discrete_8")(8)))

# If you need more than eight unique colors, simply increase the number of
# requested colors in the function call like this:
plot(prismatic::color(get_RUB_palettes(palette = "discrete_8")(16)))
```

## Plots for the discrete palettes
Last, I turned most of the unqiue colors in the RUB corporate design guide into
one awfully looking palette simply called "discrete". Avert your eyes, if at all possible! The second palette, "discrete_contrast" has four colors that form a
nice contrast to the base RUB colors, i.e. the Green and Blue.

```{r plot-discrete}
plot(prismatic::color(get_RUB_palettes(palette = "discrete")(7)))
plot(prismatic::color(get_RUB_palettes(palette = "discrete_contrast")(4)))
```

# Further Reading
The implementation is heavily indebted to Simon Jackson's great article on color palettes for ggplot2. At the moment, the University of Bochum really does not have anything like a Data Visualization Style Guide.

[Simon Jackson - Creating corporate colour palettes for ggplot2](https://drsimonj.svbtle.com/creating-corporate-colour-palettes-for-ggplot2)
[Lisa Charlotte Rost - Your Friendly Guide to Colors in Data Visualisation](https://blog.datawrapper.de/colorguide/)
[Emil Hvitfeldt's Prismatic Package](https://emilhvitfeldt.github.io/prismatic/)
[Emily Riederer's Rtistic Package](https://github.com/emilyriederer/Rtistic)
[Amy Cesal - What Are Data Visualization Style Guidelines](https://medium.com/nightingale/style-guidelines-92ebe166addc)
