---
title: "Retrieving Destatis data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Retrieving Destatis data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

This vignette will explain step-by-step how to retrieve entire table series and
their associated labels from Destatis using the package `wiesbaden`. My aim with
the approach described here is to make it easily adaptable to other use cases or 
table series. Feedback welcome!

# Wiesbaden Setup
First off, load the packages we will be using throughout this vignette.

```{r setup}
library(dplyr)
library(extrafont)
library(forcats)
library(janitor)
library(purrr)
library(RUBer)
library(tibble)
library(tidyselect)
library(wiesbaden)
```

The R package wiesbaden by Moritz Marbach provides functions to directly
retrieve data from databases maintained by the Federal Statistical Office of
Germany (Destatis) in Wiesbaden. After installing the package with
`remotes::install_github("sumtxt/wiesbaden")`, you can check 
`vignette("using-wiesbaden", package = "wiesbaden")` for more info.

In order to retrieve data from the four available databases
([Regionalstatistik](https://www.regionalstatistik.de/genesis/online),
[Genesis](https://www-genesis.destatis.de/genesis/online), [Landesdatenbank
NRW](https://www.landesdatenbank.nrw.de),
[Bildungsmonitoring](https://www.bildungsmonitoring.de)), you need to register
on the respective website. For this vignette, we use the Landesdatenbank NRW
database, so please register an account there. Once you have registered, use
`wiesbaden::save_credentials` to save your login information and
`wiesbaden::test_login` to test the connection. If you cannot establish a valid
connection, you will not be able to execute all subsequent code chunks!

```{r wiesbaden-setup, eval = FALSE}
# Save your credentials for the Landesdatenbank NRW
wiesbaden::save_credentials(
  db = "nrw",                        # Do not change
  user = "YOUR USER NAME",           # Update with your account name
  password = "YOUR PASSWORD"         # Update with your password 
)

# Test connection
wiesbaden::test_login(
   genesis = c(db = "nrw")
   )
```

# Retrieving Data
In this vignette, we are interested in tables with data on universities. On the
Landesdatenbank NRW website, the code for tables on universities is "213". So 
let us retrieve all table names with code "213":

```{r df-table, eval = FALSE}
# Retrieve all tables with code "213"
df_table <- wiesbaden::retrieve_datalist(
   tableseries= "213",
   genesis = c(db = "nrw")
   )
```

Looking at the retrieved data frame, a problem becomes apparent: the function
not only returns tables that start with code "213" but all tables with the
fragment "213" anywhere in their code. In addition to tables on universities,
this returns tables on wastewater sewage, among other things. So let us filter
the data frame to all table names whose code starts with "213":

```{r df-table-filtered, eval = FALSE}
# Filtering the tables to codes starting with "213".
df_table_filtered <- df_table %>% 
  dplyr::filter(
    stringr::str_starts(
      tablename,
      "213"
      )
  )
```

This looks much better, we now have a data frame with the names and descriptions
of 37 tables with data on universities in NRW. In the next step, we are going to
create a named vector of these table names using `purrr::set_names`. We then
download the data for all table names in this vector and put them in one big
data frame using a combination of `purrr::map_dfr` and
`wiesbaden::retrieve_data`. You will be downloading a lot of data, so this step
may take a while!

```{r get-data, eval = FALSE}
# Get all table names and turn them into a named vector
table_names <- purrr::set_names(
  x = as.character(
    df_table_filtered[["tablename"]]
  ),
  nm = as.character(
      df_table_filtered[["tablename"]]
      )
  )

# Retrieve data for every table name and bind them into one data frame. Note
# that this may take a while!
df <- purrr::map_dfr(
  table_names,
  wiesbaden::retrieve_data,
  genesis = c(db = "nrw"),
  .id = "tablename"
  )
```

For each table, we now download the available metadata with
`wiesbaden::retrieve_metadata` and put them into one data frame using
`purrr::map_dfr`. The metadata describes the different variables used in each
table. The metadata consists of four columns: tablename, (variable) name,
(variable) description and (variable) unit. We then create a character vector of
the unique variable names for all tables.

```{r get-metadata, eval = FALSE}
# Retrieve metadata for every table name and bind them into one data frame
df_metadata <- purrr::map_dfr(
  table_names,
  wiesbaden::retrieve_metadata,
  genesis = c(db = "nrw"),
  .id = "tablename"
)

# Create character vector of all unique variable names in all tables
variables_unique <- unique(
  as.character(
    df_metadata[["name"]]
    )
  )
```

We now have a vector of 30 unique variables that are used in the table series
"213". The data tables use value keys and not the actual value labels (e.g. the
variable "SX-A01" has the two value keys "SX-OPSCHUL1" and "SX-OPSCHUL2" with
the labels "Öffentlich" (*Public*) and "Privat" (*private*) stored in separate
key tables). We use a combination of `purrr::map` and
`wiesbaden::retrieve_valuelabel` to retrieve all value labels for our data
tables in a list. We then do some data cleaning using `purrr::discard`.

```{r get-value-labels, eval = FALSE}
# Retrieve value labels with wiesbaden::retrieve_valuelabel
list_value_labels <- purrr::map(
  variables_unique,
  wiesbaden::retrieve_valuelabel,
  genesis = c(db = "nrw")
  ) 

# Variables with no labels return a list with the character "No results found."
# We use purrr::discard to drop all these empty list elements of type character.
list_value_labels_filtered <- list_value_labels %>% 
  purrr::discard(
    ~ is.character(.x)
  )
```

Unlike the downloaded data and metadata, which we have put into one data frame
each, the value labels are put into a list of data frames. To retrieve a single
element from that list, i.e. a single data frame of value labels, use double
square brackets like this `list_value_labels_filtered[[1]]`.

# Data Wrangling
We now have all the data required and have done some minimal data cleaning and
filtering. In this section, we clean the data some more and bring it all
together in one data frame.

The Destatis data uses capitalized column names and column names containing
hyphens. We automatically change this to snake case and replace special
characters for all column names using `janitor::clean_names`. The data also
includes three types of columns that do not seem to be documented and which
serve no apparent purpose. Specifically, the undocumented columns ending in
"qual" (the only value being "e"), "err" (always 0) and "lock" (always NA) are
of no value to us, so we discard them from the data frame.

```{r wrangle-data, eval = FALSE}
# Clean all column names (e.g. replacing hyphens with underscores) and make them
# use snake case.
# Discard columns ending in "qual", "err", or "lock".
df_cleaned <- df %>%
  janitor::clean_names(
    case = "snake"
  ) %>% 
  dplyr::select(
    -ends_with(
      c(
        "qual",
        "err", 
        "lock"
        )
      )
    )
```

We clean the list of value labels by turning all factors into character vectors
using `dplyr::mutate_all`. Factors can be very useful, but I think it is easiest
to use them selectively and only set them with a specific use case in mind. We
then apply `janitor::clean_names` to the list of data frames containing value
labels. As we have also applied this function to the data frame with the actual
data, this makes the column names between the different tables consistent.

We want to combine all value labels with the data frame containing the actual
data. Unfortunately, each data frame with value labels has a column 
"description". We need to make this column unique across the list of data frames
containing value labels, so that we know which value key the description applies 
to. First, we use `purrr::map_chr` to generate a character vector of the 
descriptions. This is then used as the input to another function, where we 
rename all the description columns to "description_" plus "key_id".

```{r wrangle-value-labels, eval = FALSE}
# Turn factors into character vectors and clean all column names with janitor
list_value_labels_wrangled <- list_value_labels_filtered %>% 
  purrr::map(
    dplyr::mutate_all,
    as.character
  ) %>% 
  purrr::map(
    janitor::clean_names,
    case = "snake"
  )

# Generate a character vector of unique description column names
description_names <- purrr::map_chr(
  purrr::map(
    list_value_labels_wrangled,
    ~ paste0(names(.x), "_description")
  ),
  1
)

# Rename the existing description column names to make them unique
list_value_labels_unique <- purrr::map2(
  list_value_labels_wrangled,
  description_names,
  ~dplyr::rename(
    .x, !!.y := description
    )
)
```

We are now ready to combine the data with the corresponding labels. Before,
though, we will retrieve another set of labels, describing the contents of each
table. As with the value labels, we rename the generic "description" column to
"tablename_description" to make it unique.

For combining tables that share a mutual key or mutal set of keys, we use
`dplyr::left_join`. Because we have so many individual tables and keys, doing 
this manually would be tedious. To avoid this manual process, we use 
`purrr::reduce` on a list of all data frames we want to join. So before applying 
`purrr::reduce`, we put the data frame with the actual data, the data frame with
the table descriptions, and the list of data frames containing value labels into 
one big list.

```{r join-data-to-labels, eval = FALSE}
# Create data frame with table name and table description
df_table_descriptions <- df_table_filtered %>% 
  dplyr::mutate_all(
    as.character
  ) %>% 
  dplyr::select(
   tablename,
   tablename_description = description
  )

# Combine actual data with the data frame of table names and the list of data 
# frames containing value labels into one list.
df_list_tmp <- append(
  list(df_cleaned),
  list(df_table_descriptions)
)

df_list <- append(
  df_list_tmp,
  list_value_labels_unique
)

# Join data with all the value label data frames to create one big data frame
df_joined <- purrr::reduce(
  df_list,
  dplyr::left_join
)
```

We also sort the columns (not the rows!) alphabetically, so that the
descriptions come immediately after the respective key column. The first two
columns are always the table code and the table description. We also turn the
data frame into a tibble, for better print defaults among other benefits.

```{r finalize-data, eval = FALSE}
# Order columns alphabetically, put columns tablename and tablename_description 
# first, turn data frame into tibble.
db_nrw_213 <- df_joined %>% 
  dplyr::select(
    sort(
      tidyselect::peek_vars()
      )
  ) %>% 
  dplyr::select(
    tablename,
    tablename_description,
    everything()
  ) %>% 
  tibble::as_tibble()
```

The resulting data frame, `db_nrw_213`, actually comes bundled with the `RUBer`
package. We use `usethis::use_data` to handle this step for us. We have now 
added 37 tables of university related data to our package!

```{r save-data, eval = FALSE}
usethis::use_data(
  db_nrw_213,
  overwrite = TRUE,
  version = 3
  )

load(
  "D:\\Git\\db_nrw_213.rda"
)
```

# Working with the Destatis data
We usually do not want to work with the entire data frame at once. The RUBer
package provides the function `filter_destatis_code`, which allows you to filter
the data based on a table code. It also discards all columns that only consist
of NA after filtering. As an example, let us create a data frame consisting of
all data for table code "21311LS001A".

```{r filter-destatis-code, eval = FALSE}
# Get all rows and columns for table code 21311LS001A
df_21311LS001A <- filter_destatis_code(
  df = db_nrw_213,
  tablename = "21311LS001A"
  )
```

The table description says that this table contains "Studierende / Studenten,
Land, Hochschulen, Studienfach, Fachsemester, Angestrebte Abschlussprüfung,
Semester". So essentially it is a table of students for each degree program,
term, and university. We will do some basic analysis and plotting using this 
data.

First, let us find out which five universities had the highest overall number of
students with the help of `forcats::fct_lump`.

```{r 21311LS001A-plot1, eval = FALSE}
df_plot <- df_21311LS001A %>% 
  dplyr::filter(
    semest == "WS 2017/18"
  ) %>% 
  dplyr::mutate(
    bilhs1_description = forcats::fct_lump_n(
      bilhs1_description,
      n = 5,
      w = bil002_val,
      other_level = "Other NRW universities"
      ),
    bilap1_description = forcats::fct_lump_n(
      bilap1_description,
      n = 5,
      w = bil002_val,
      other_level = "Other degrees"
      )
    ) %>% 
  dplyr::filter(
    bilhs1_description != "Other NRW universities"
  )

rub_plot_type_1(
  df = df_plot,
  x_var = bilhs1_description,
  y_var = bil002_val,
  fill_var = bilap1_description,
  y_axis_label = "Students in winter term 2017/18",
  caption = "Landesdatenbank NRW, Table Code 21311LS001A (license dl-de/by-2-0)",
  caption_prefix = "Source:",
  filter_cutoff = 0.06
#  ,fill_reverse = TRUE
)
```

The Fernuniversität Hagen, the TH Aachen, the University of Bochum, the
University of Köln and the University of Münster had the highest student count
in the winter term 2017/18. Maybe we are interested more specifically in the 
relative importance of each degree type at each university. We can use 
`rub_plot_type_2` for that:

```{r 21311LS001A-plot2, eval = FALSE}
rub_plot_type_2(
  df = df_plot,
  x_var = bilhs1_description,
  y_var = bil002_val,
  y_axis_label = "Students in winter term 2017/18",
  fill_var = bilap1_description,
  caption = "Landesdatenbank NRW, Table Code 21311LS001A (license dl-de/by-2-0)",
  caption_prefix = "Source:",
  filter_cutoff = 0.06
#  ,fill_reverse = TRUE
)
```

# Data license for the Destatis data

Data obtained from Destatis is available under the Data licence Germany –
attribution – version 2.0 (dl-de/by-2-0). The license text is available at
[www.govdata.de/dl-de/by-2-0](www.govdata.de/dl-de/by-2-0). The data in this
vignette was obtained from the Landesdatenbank NRW using the code "213" for
"Hochschulen". The license requests a reference to the dataset using the URI. To
the best of my knowledge, the Landesdatenbank NRW does not yet provide Uniform
Resource Identifiers (URI) to identify these tables.


# Data Analysis

```{r dev, eval = FALSE}
debugonce(rub_plot_type_1)
devtools::document()
devtools::load_all()
```
