---
title: "Retrieving Destatis data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Retrieving Destatis data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

This vignette will explain step-by-step how to retrieve entire table series and
their associated labels from Destatis using the package `wiesbaden`. My aim with
the approach described here is to make it easily adaptable to other use cases or 
table series. Feedback welcome!

# Wiesbaden Setup
First off, load the packages we will be using throughout this vignette.

```{r libraries}
library(dplyr)
library(extrafont)
library(forcats)
library(httptest)
library(janitor)
library(purrr)
library(RUBer)
library(tibble)
library(tidyselect)
library(wiesbaden)
```

```{r httptest, include = FALSE}
api_root <- "https://www.landesdatenbank.nrw.de/ldbnrwws/services/"

# Makes sure that the user and password information is not stored in the cached
# HTTP responses.
# Shorten URLs
set_redactor(function (response) {
    response %>%
      gsub_response(
        "kennung=[^&]+",
        "kennung=REDACTED"
      ) %>%
      gsub_response(
        "passwort=[^&]+",
        "passwort=REDACTED"
      ) %>% 
      gsub_response(
        api_root,
        "",
        fixed = TRUE
      )
})

# Makes sure that the user and password information is not stored in the cached
# HTTP responses.
# Shorten URLs
set_requester(function (request) {
    request %>%
      gsub_request(
        "kennung=[^&]+",
        "kennung=REDACTED"
      ) %>%
      gsub_request(
        "passwort=[^&]+",
        "passwort=REDACTED"
      ) %>%   
      gsub_request(
        api_root,
        "",
        fixed = TRUE
      )
})

httptest::start_vignette("wiesbaden")
```

The R package `wiesbaden` by Moritz Marbach provides functions to directly
retrieve data from databases maintained by the Federal Statistical Office of
Germany (Destatis) in Wiesbaden. After installing the package with
`remotes::install_github("sumtxt/wiesbaden")`, you can check 
`vignette("using-wiesbaden", package = "wiesbaden")` for more info.

In order to retrieve data from the four available databases
([Bildungsmonitoring](https://www.bildungsmonitoring.de), [Genesis](https://www-genesis.destatis.de/genesis/online), [Landesdatenbank
NRW](https://www.landesdatenbank.nrw.de), [Regionalstatistik](https://www.regionalstatistik.de/genesis/online)), you need to register
on the respective website. For this vignette, we use the Landesdatenbank NRW
database, so please register an account there. Once you have registered, use
`wiesbaden::save_credentials` to save your login information and
`wiesbaden::test_login` to test the connection. If you cannot establish a valid
connection, you will not be able to execute all subsequent code chunks!

```{r wiesbaden-setup, eval = FALSE}
# Save your credentials for the Landesdatenbank NRW
wiesbaden::save_credentials(
  db = "nrw",                        # Do not change
  user = "YOUR USER NAME",           # Update with your account name
  password = "YOUR PASSWORD"         # Update with your password 
)

# Test connection
wiesbaden::test_login(
   genesis = c(db = "nrw")
   )
```

# Retrieving Data
In this vignette, we are interested in tables with data on universities. On the
Landesdatenbank NRW website, the code for tables on universities is "213". So 
let us retrieve all table names with code "213" and turn the resulting data 
frame into a tibble for prettier printing.

```{r df-table}
# Retrieve all tables with code "213" and turn data frame into tibble
df_table <- wiesbaden::retrieve_datalist(
   tableseries= "21311*",
   genesis = c(db = "nrw")
   ) %>% 
  tibble::as_tibble()

df_table
```

We now have a data frame with the names and descriptions of 37 tables with data on universities in NRW. In the next step, we are going to create a named vector of these table names using `purrr::set_names`. We then download the data for all table names in this vector and put them in one big data frame using a combination of `purrr::map_dfr` and `wiesbaden::retrieve_data`. You will be downloading a lot of data, so this step may take a while!

```{r get-data, message = FALSE}
# Get all table names and turn them into a named vector
table_names <- purrr::set_names(
  x = as.character(
    df_table[["tablename"]]
  ),
  nm = as.character(
      df_table[["tablename"]]
      )
  )

table_names

# Retrieve data for every table name and bind them into one data frame. Note
# that this may take a while!
df <- purrr::map_dfr(
  table_names,
  wiesbaden::retrieve_data,
  genesis = c(db = "nrw"),
  .id = "tablename"
  ) %>% 
  tibble::as_tibble()

df
```

For each table, we now download the available metadata with
`wiesbaden::retrieve_metadata` and put them into one data frame using
`purrr::map_dfr`. The metadata describes the different variables used in each
table. The metadata consists of four columns: tablename, (variable) name,
(variable) description and (variable) unit. We then create a character vector of
the unique variable names for all tables.

```{r get-metadata}
# Retrieve metadata for every table name and bind them into one data frame
df_metadata <- purrr::map_dfr(
  table_names,
  wiesbaden::retrieve_metadata,
  genesis = c(db = "nrw"),
  .id = "tablename"
) %>% 
  tibble::as_tibble()

df_metadata

# Create character vector of all unique variable names in all tables
variables_unique <- unique(
  as.character(
    df_metadata[["name"]]
    )
  )

variables_unique
```

We now have a vector of 30 unique variables that are used in the table series
"213". The data tables use value keys and not the actual value labels (e.g. the
variable "SX-A01" has the two value keys "SX-OPSCHUL1" and "SX-OPSCHUL2" with
the labels "Öffentlich" (*Public*) and "Privat" (*private*) stored in separate
key tables). We use a combination of `purrr::map` and
`wiesbaden::retrieve_valuelabel` to retrieve all value labels for our data
tables in a list.

```{r get-value-labels}
# Retrieve value labels with wiesbaden::retrieve_valuelabel
list_value_labels <- purrr::map(
  variables_unique,
  wiesbaden::retrieve_valuelabel,
  genesis = c(db = "nrw")
  )

str(list_value_labels)
```

Unlike the downloaded data and metadata, which we have put into one data frame
each, the value labels are put into a list of data frames. To retrieve a single
element from that list, i.e. a single data frame of value labels, use double
square brackets like this `list_value_labels_filtered[[1]]`.

# Data Wrangling
We now have all the data required and have done some minimal data cleaning and
filtering. In this section, we clean the data some more and bring it all
together in one data frame.

The Destatis data uses capitalized column names and column names containing
hyphens. We automatically change this to snake case and replace special
characters for all column names using `janitor::clean_names`. The data also
includes three types of columns that do not seem to be documented and which
serve no apparent purpose. Specifically, the undocumented columns ending in
"qual" (the only value being "e"), "err" (always 0) and "lock" (always NA) are
of no value to us, so we discard them from the data frame.

```{r wrangle-data}
# Clean all column names (e.g. replacing hyphens with underscores) and make them
# use snake case.
# Discard columns ending in "qual", "err", or "lock".
df_cleaned <- df %>%
  janitor::clean_names(
    case = "snake"
  ) %>% 
  dplyr::select(
    -ends_with(
      c(
        "qual",
        "err", 
        "lock"
        )
      )
    )

df_cleaned
```

In our list of value labels, some elements contain the string "No results 
found". We drop these empty elements using `purrr::discard`. We
then apply `janitor::clean_names` to the list of data frames containing value
labels. As we have also applied this function to the data frame with the actual
data, this makes the column names between the different tables consistent.

Next, we want to 
combine all value labels with the data frame containing the actual data. 
Unfortunately, each data frame with value labels has a column "description". We 
need to make this column unique across the list of data frames containing value 
labels, so that we know which value key the description applies to. First, we 
use `purrr::map_chr` to generate a character vector of the descriptions. This is
then used as the input to another function, where we rename all the description 
columns to "description_" plus "key_id".

```{r wrangle-value-labels}
# Variables with no labels return a list with the character "No results found."
# We use purrr::discard to drop all these empty list elements of type character.
list_value_labels_filtered <- purrr::discard(
  list_value_labels,
  ~ is.character(.x)
  )

str(list_value_labels_filtered)

str(list_value_labels_filtered[[1]][["description"]])

# Clean all column names with janitor
list_value_labels_wrangled <- list_value_labels_filtered %>% 
  purrr::map(
    janitor::clean_names,
    case = "snake"
  )

str(list_value_labels_wrangled)

# Generate a character vector of unique description column names
description_names <- purrr::map_chr(
  purrr::map(
    list_value_labels_wrangled,
    ~ paste0(names(.x), "_description")
  ),
  1
)

description_names

# Rename the existing description column names to make them unique
list_value_labels_unique <- purrr::map2(
  list_value_labels_wrangled,
  description_names,
  ~dplyr::rename(
    .x,
    !!.y := description
    )
)

str(list_value_labels_unique)
```

We are now ready to combine the data with the corresponding labels. Before,
though, we will retrieve another set of labels, describing the contents of each
table. As with the value labels, we rename the generic "description" column to
"tablename_description" to make it unique.

For combining tables that share a mutual key or mutal set of keys, we use
`dplyr::left_join`. Because we have so many individual tables and keys, doing 
this manually would be tedious. To avoid this manual process, we use 
`purrr::reduce` on a list of all data frames we want to join. So before applying 
`purrr::reduce`, we put the data frame with the actual data, the data frame with
the table descriptions, and the list of data frames containing value labels into 
one big list.

```{r join-data-to-labels}
# Create data frame with table name and table description
df_table_descriptions <- df_table %>% 
  dplyr::rename(
   tablename_description = description
  )

df_table_descriptions

# Combine actual data with the data frame of table names and the list of data 
# frames containing value labels into one list.
df_list_tmp <- append(
  list(df_cleaned),
  list(df_table_descriptions)
)

df_list <- append(
  df_list_tmp,
  list_value_labels_unique
)

str(df_list)

# Join data with all the value label data frames to create one big data frame
df_joined <- purrr::reduce(
  df_list,
  dplyr::left_join
) %>% 
  tibble::as_tibble()

df_joined
```

We also sort the columns (not the rows!) alphabetically, so that the
descriptions come immediately after the respective key column. The first two
columns are always the table code and the table description.

```{r finalize-data}
# Order columns alphabetically, put columns tablename and tablename_description 
# first, put value columns last.
db_nrw_213 <- df_joined %>% 
  dplyr::select(
    sort(
      tidyselect::peek_vars()
      )
  ) %>% 
  dplyr::relocate(
    tablename,
    tablename_description,
    !(ends_with("_val")),
    ends_with("val")
    )

db_nrw_213
```

# Working with the Destatis data

We usually do not want to work with the entire data frame at once. The RUBer
package provides the function `filter_destatis_code`, which allows you to filter
the data based on a table code. It also discards all columns that only consist
of NA after filtering. As an example, let us create a data frame consisting of
all data for table code "21311LS001A".

```{r filter-destatis-code}
# Get all rows and columns for table code 21311LS001A
df_21311LS001A <- filter_destatis_code(
  df = db_nrw_213,
  tablename = "21311LS001A"
  )
```

The table description says that this table contains "Studierende / Studenten,
Land, Hochschulen, Studienfach, Fachsemester, Angestrebte Abschlussprüfung,
Semester". So essentially it is a table of students for each degree program,
term, and university. We will do some basic analysis and plotting using this 
data.

First, let us find out which five universities had the highest overall number of
students with the help of `forcats::fct_lump_n`.

```{r 21311LS001A-plot1, fig.cap = "Students at the five biggest universities in NRW for the winter term 2017/18 for each degree (Vertical stacked bar chart)"}
# Create data frame with plotting data, filtering on the winter term 2017/18,
# focusing on the top five universities in terms of student count and keeping
# the top five degree types, in addition to a lumped category "other degrees".
df_plot <- df_21311LS001A %>% 
  dplyr::filter(
    semest == "WS 2017/18"
  ) %>%   
  dplyr::mutate(
    bilhs1_description = forcats::fct_lump_n(
      bilhs1_description,
      n = 5,
      w = bil002_val,
      other_level = "Other NRW universities"
      )
  ) %>%
  dplyr::filter(
    bilhs1_description != "Other NRW universities"
  ) %>% 
  dplyr::mutate(
    bilap1 = forcats::fct_lump_n(
      bilap1,
      n = 5,
      w = bil002_val,
      other_level = "Other degrees"
      ),
    bilap1_description = forcats::fct_lump_n(
      bilap1_description,
      n = 5,
      w = bil002_val,
      other_level = "Other degrees"
      )
    )

rub_plot_type_1(
  df = df_plot,
  x_var = bilhs1_description,
  y_var = bil002_val,
  fill_var = bilap1_description,
  y_axis_label = "Students in winter term 2017/18",
  caption = "Landesdatenbank NRW, Table Code 21311LS001A (license dl-de/by-2-0)",
  caption_prefix = "Source:",
  filter_cutoff = 0.06
#  ,fill_reverse = TRUE
)
```

The Fernuniversität Hagen, the TH Aachen, the University of Bochum, the
University of Köln and the University of Münster had the highest student count
in the winter term 2017/18. Maybe we are interested more specifically in the 
relative importance of each degree type at each university. We can use 
`rub_plot_type_2` for that:

```{r 21311LS001A-plot2, fig.cap = "Students at the five biggest universities in NRW for the winter term 2017/18 for each degree (Vertical stacked bar chart scaled to 100 percent)"}
rub_plot_type_2(
  df = df_plot,
  x_var = bilhs1_description,
  y_var = bil002_val,
  y_axis_label = "Students in winter term 2017/18",
  fill_var = bilap1_description,
  caption = "Landesdatenbank NRW, Table Code 21311LS001A (license dl-de/by-2-0)",
  caption_prefix = "Source:",
  filter_cutoff = 0.05
# ,fill_reverse = TRUE
# ,palette_reverse = TRUE
)
```

If we wanted to examine how the relative importance of each degree has developed
over time, we could switch to `rub_plot_type_3` (a rotated version of type 2),
use the term as y-variable and the university name as facet variable.

```{r 21311LS001A-plot3, fig.cap = "Students at the five biggest universities in NRW from 2013 to 2018 for each degree (Vertical stacked bar chart scaled to 100 percent)"}
# Create data frame with plotting data, filtering on the winter terms 2013-2018,
# focusing on the top five universities in terms of student count and keeping
# the top five degree types, in addition to a lumped category "other degrees".
df_plot_t3 <- df_21311LS001A %>% 
  dplyr::filter(
    semest %in% c(
      "WS 2017/18",
      "WS 2016/17",
      "WS 2015/16",
      "WS 2014/15",
      "WS 2013/14"
      )
    ) %>% 
  dplyr::mutate(
    bilhs1_description = forcats::fct_lump_n(
      bilhs1_description,
      n = 5,
      w = bil002_val,
      other_level = "Other NRW universities"
      )
  ) %>%
  dplyr::filter(
    bilhs1_description != "Other NRW universities"
  ) %>% 
  dplyr::mutate(
    bilap1 = forcats::fct_lump_n(
      bilap1,
      n = 5,
      w = bil002_val,
      other_level = "Other degrees"
      ),
    bilap1_description = forcats::fct_lump_n(
      bilap1_description,
      n = 5,
      w = bil002_val,
      other_level = "Other degrees"
      )
    )

rub_plot_type_3(
  df = df_plot_t3,
  x_var = bil002_val,
  y_var = semest,
  facet_var = bilhs1_description,
  fill_var = bilap1_description,
  caption = "Landesdatenbank NRW, Table Code 21311LS001A (license dl-de/by-2-0)",
  caption_prefix = "Source:",
  filter_cutoff = 0.06
  ,fill_reverse = FALSE
)
```

# Data license for the Destatis data

Data obtained from Destatis is available under the [Data licence Germany –
attribution – version 2.0 (dl-de/by-2-0)](www.govdata.de/dl-de/by-2-0). The data
in this vignette was obtained from the Landesdatenbank NRW using the code "213"
for "Hochschulen". The license requests a reference to the dataset using the
URI. To the best of my knowledge, the Landesdatenbank NRW does not yet provide
Uniform Resource Identifiers (URI) to identify these tables.

```{r httptest-end, include=FALSE}
httptest::end_vignette()
```
